---
title: "Proof of Concept - GloVe Distributions"
output: github_document
---
  
This script aims to prove the following concepts:

* The average GloVe position of words in a sentence is a meaningful way of 
representing the topic, theme or meaning of the sentence.
* The distribution of these sentence-based GloVe positions can be approximated
by a standard statistical distribution, and similar impact case studies should
have significant overlap in their distributions.
* Individual sentences which are outliers from the overall distribution can be
identified programmatically.

## Packages Required

This script will use **text2vec** for the GloVe algorithm, **tibble** for use of
the data_frame, **stringr** for string operations, **tidytext** for managing 
the dataset, and **ggplot2** for plotting.

```{r setup, message=FALSE}
library(text2vec)
library(stringr)
library(tibble)
library(tidytext)
library(ggplot2)
```

## Data Import and Tidy Text

The data has been downloaded previously using the **refimpact** package, and has
been saved locally to speed up analysis. The `unnest_tokens()` function from
the **tidytext** package is used to clean up the text and break it into 
sentences.

```{r import}
ref <- readRDS("~/ref_data.rds")
ref <- ref[,c("CaseStudyId", "UOA", "ImpactType",
              "Institution", "Title", "ImpactDetails")]
tidy_ref <- unnest_tokens(ref, text, ImpactDetails, "sentences")
glimpse(tidy_ref)
```

## GloVe

GloVe is effectively a dimensionality reduction technique for bag-of-words text
representation, where each word is mapped into a position in m-dimensional space
based on its observed context in the corpus. Due to the importance of context,
it needs to be trained on the original unbroken impact studies rather than the
tidy text (which has been broken into sentences).

For the initial investigation of the nature of the distributions generated, the
GloVe model will only use 2 dimensions. For the final model the number of 
dimensions will likely be 50.

```{r glove-2d, message=FALSE}
prepare_text <- function(x) {
  x %>% 
  str_to_lower %>% 
  str_replace_all("[^[:alnum:]]", " ") %>% 
  str_replace_all("\\s+", " ")
}

glove_input <- prepare_text(ref$ImpactDetails)
tokens <- space_tokenizer(glove_input)
it <- itoken(tokens)
vocab <- create_vocabulary(it)
vocab <- prune_vocabulary(vocab, term_count_min = 5L)
vectorizer <- vocab_vectorizer(vocab, grow_dtm = FALSE, skip_grams_window = 5L)
tcm <- create_tcm(it, vectorizer)

glove <- GlobalVectors$new(word_vectors_size = 2,
                           vocabulary = vocab,
                           x_max = 10L,
                           #lambda = 1e-5,
                           shuffle=T)
glove$fit(tcm, n_iter = 50)

word_vectors <- glove$get_word_vectors()
```

This GloVe model has now mapped each word from the corpus into a two dimensional
space where (in practice) related words are grouped together and the vector 
differences between points have meaning. The standard results published in the
**word2vec** and **GloVe** papers are:

* king - man + woman = queen
* berlin - germany + italy = rome

We can take a look at some of these mappings by inspecting the word_vectors 
data frame.

```{r inspect}
word_vectors[c("cat", "dog", "research", "impact"),]
```

We can also look at how these words are distributed across the two dimensional
space.

```{r plot-2d}
words <- as.data.frame(word_vectors)
names(words) <- c("dim_1", "dim_2")
ggplot(words) +
  geom_point(aes(x=dim_1, y=dim_2), alpha = 0.05) +
  geom_density2d(aes(x=dim_1, y=dim_2))
```

This is good! They sort of resemble a normal distribution with a bit of noise, 
and this may well improve as the dimensionality of the GloVe model increases. 
But why assume? We can test this by fitting a 50 dimensional GloVe model and
then plotting any two dimensions (noting that all dimensions are equally
important in GloVe, unlike other dimensionality reduction techniques like PCA).

```{r glove-50d, message=FALSE}
glove2 <- GlobalVectors$new(word_vectors_size = 50,
                           vocabulary = vocab,
                           x_max = 10L,
                           #lambda = 1e-5,
                           shuffle=T)
glove2$fit(tcm, n_iter = 50)

word_vectors2 <- glove2$get_word_vectors()
words2 <- as.data.frame(word_vectors2[,c(1,2)])
names(words2) <- c("dim_1", "dim_2")
ggplot(words2) +
  geom_point(aes(x=dim_1, y=dim_2), alpha = 0.05) +
  geom_density2d(aes(x=dim_1, y=dim_2))
```

This looks like the assumption of multivariate normality might be okay! Woohoo! 
While we're here we might as well test it formally using the **MVN** package.
The package doesn't seem to deal so well with large datasets, so we'll test
normality with a sample.

```{r multi-normality}
library(MVN)
mardiaTest(word_vectors2[sample(1:nrow(word_vectors2), 5000),])
hzTest(word_vectors2[sample(1:nrow(word_vectors2), 5000),])
roystonTest(word_vectors2[sample(1:nrow(word_vectors2), 2000),])
```

These are all pretty resounding "no" results from the formal tests, however it 
is quite hard to find a large, 50-dimensional dataset for which the multivariate
normality assumption holds. The important thing is that we can use the 
almost-multivariate-normal assumption to find outliers, but this won't be 
necessary until we're looking at whole sentences.

Before we move on, we might as well check that similar words are near each 
other, and different words are far apart. We can do this visually using the 
2 dimensional GloVe model.

```{r check-1}
word_vectors[c("kidney", "liver", "institute", "university"),]
```

It looks like it works pretty well, even in 2 dimensions! To be extra certain
we can look at the euclidian distances between the words in the 50 dimensional
GloVe model.

```{r check-2}
dist(word_vectors2[c("kidney", "liver", "institute", "university"),],
     method = "euclidian")
```

This is as expected - kidney and liver are the closest two words, then 
institute and university are the next closest pairing. The 50-dimensonal 
representation is doing a reasonable job of separating out the words.

## GloVe Representation of Sentences

To be continued...
