---
title: "Proof of Concept - GloVe Distributions"
output: github_document
---
  
This script aims to prove the following concepts:

* The average GloVe position of words in a sentence is a meaningful way of 
representing the topic, theme or meaning of the sentence.
* The distribution of these sentence-based GloVe positions can be approximated
by a standard statistical distribution, and similar impact case studies should
have significant overlap in their distributions.
* Individual sentences which are outliers from the overall distribution can be
identified programmatically.

## Packages Required

This script will use **text2vec** for the GloVe algorithm, **tibble** for use of
the data_frame, **stringr** for string operations, **tidytext** for managing 
the dataset, and **ggplot2** for plotting.

```{r setup, message=FALSE}
library(text2vec)
library(stringr)
library(tibble)
library(tidytext)
library(ggplot2)
```

## Data Import and Tidy Text

The data has been downloaded previously using the **refimpact** package, and has
been saved locally to speed up analysis. The `unnest_tokens()` function from
the **tidytext** package is used to clean up the text and break it into 
sentences.

```{r import}
ref <- readRDS("~/ref_data.rds")
ref <- ref[,c("CaseStudyId", "UOA", "ImpactType",
              "Institution", "Title", "ImpactDetails")]
tidy_ref <- unnest_tokens(ref, text, ImpactDetails, "sentences")
glimpse(tidy_ref)
```

## GloVe

GloVe is effectively a dimensionality reduction technique for bag-of-words text
representation, where each word is mapped into a position in m-dimensional space
based on its observed context in the corpus. Due to the importance of context,
it needs to be trained on the original unbroken impact studies rather than the
tidy text (which has been broken into sentences).

For the initial investigation of the nature of the distributions generated, the
GloVe model will only use 2 dimensions. For the final model the number of 
dimensions will likely be 50.

```{r glove-2d, message=FALSE}
prepare_text <- function(x) {
  x %>% 
  str_to_lower %>% 
  str_replace_all("[^[:alnum:]]", " ") %>% 
  str_replace_all("\\s+", " ")
}

glove_input <- prepare_text(ref$ImpactDetails)
tokens <- word_tokenizer(glove_input)
it <- itoken(tokens)
vocab <- create_vocabulary(it)
vocab <- prune_vocabulary(vocab, term_count_min = 5L)
vectorizer <- vocab_vectorizer(vocab, grow_dtm = FALSE, skip_grams_window = 5L)
tcm <- create_tcm(it, vectorizer)

glove <- GlobalVectors$new(word_vectors_size = 2,
                           vocabulary = vocab,
                           x_max = 10L,
                           #lambda = 1e-5,
                           shuffle=T)
glove$fit(tcm, n_iter = 50)

word_vectors <- glove$get_word_vectors()
```

This GloVe model has now mapped each word from the corpus into a two dimensional
space where (in practice) related words are grouped together and the vector 
differences between points have meaning. The standard results published in the
**word2vec** and **GloVe** papers are:

* king - man + woman = queen
* berlin - germany + italy = rome

We can take a look at some of these mappings by inspecting the word_vectors 
data frame.

```{r inspect}
word_vectors[c("cat", "dog", "research", "impact"),]
```

We can also look at how these words are distributed across the two dimensional
space.

```{r plot-2d}
words <- as.data.frame(word_vectors)
names(words) <- c("dim_1", "dim_2")
ggplot(words) +
  geom_point(aes(x=dim_1, y=dim_2), alpha = 0.05) +
  geom_density2d(aes(x=dim_1, y=dim_2))
```

This is good! They sort of resemble a normal distribution with a bit of noise, 
and this may well improve as the dimensionality of the GloVe model increases. 
But why assume? We can test this by fitting a 50 dimensional GloVe model and
then plotting any two dimensions (noting that all dimensions are equally
important in GloVe, unlike other dimensionality reduction techniques like PCA).

```{r glove-50d, message=FALSE}
glove2 <- GlobalVectors$new(word_vectors_size = 50,
                           vocabulary = vocab,
                           x_max = 10L,
                           lambda = 1e-5,
                           shuffle=T)
glove2$fit(tcm, n_iter = 50)

word_vectors2 <- glove2$get_word_vectors()
words2 <- as.data.frame(word_vectors2[,c(1,2)])
names(words2) <- c("dim_1", "dim_2")
ggplot(words2) +
  geom_point(aes(x=dim_1, y=dim_2), alpha = 0.05) +
  geom_density2d(aes(x=dim_1, y=dim_2))
```

This looks like the assumption of multivariate normality might be okay! Woohoo! 
While we're here we might as well test it formally using the **MVN** package.
The package doesn't seem to deal so well with large datasets, so we'll test
normality with a sample.

```{r multi-normality}
library(MVN)
mardiaTest(word_vectors2[sample(1:nrow(word_vectors2), 5000),])
hzTest(word_vectors2[sample(1:nrow(word_vectors2), 5000),])
roystonTest(word_vectors2[sample(1:nrow(word_vectors2), 2000),])
```

These are all pretty resounding "no" results from the formal tests, however it 
is quite hard to find a large, 50-dimensional dataset for which the multivariate
normality assumption holds. The important thing is that we can use the 
almost-multivariate-normal assumption to find outliers, but this won't be 
necessary until we're looking at whole sentences.

Before we move on, we might as well check that similar words are near each 
other, and different words are far apart. We can do this visually using the 
2 dimensional GloVe model.

```{r check-1}
word_vectors[c("kidney", "bladder", "institute", "university"),]
```

It looks like it works pretty well, even in 2 dimensions! To be extra certain
we can look at similar words in the 50 dimensional GloVe model.

```{r check-2}
get_nearest <- function(x) {
  sim2(x = word_vectors2, y = word_vectors2[x,,drop=F], 
       method = "cosine", norm = "l2")[,1] %>% 
    sort(decreasing = TRUE) %>% head(5)
}
get_nearest("kidney")
get_nearest("bladder")
get_nearest("institute")
get_nearest("university")
```

This is looking promising - similar words are near each other in the 50D 
representation, which suggests it is doing a reasonable job of separating out 
the words.

The final thing we can do to assess the GloVe representation is to test the word
relationships using a standard test set. This is unlikely to perform very well
on this dataset as it is quite specific, however it should provide a reasonable
basis for comparison when tuning the GloVe model later in the project.

```{r}
questions_file <- '~/impactface/Data/questions-words.txt'
qlst <- prepare_analogy_questions(questions_file,
                                  rownames(word_vectors2),
                                  verbose = T)
res <- check_analogy_accuracy(questions_list = qlst, 
                              m_word_vectors = word_vectors2,
                              verbose = T)
```

The overall accuracy isn't hugely important in this context, and it will be 
improved upon when the model is tuned.

## GloVe Representation of Sentences

There is currently no universally accepted method for combining GloVe (or
word2vec) vectors into sentences. Some approaches involve training a new model
to look at sentence context, however this requires significantly more training
data than I have access to. Other approaches include either:

1. taking the average vector for each of the words in the sentence, or
1. applying the tf-idf weighted average vector for each of the words in the 
sentence. 

Given that the second option should up-weight important words and down-weight 
unimportant words, it seems like a good option. We can implement this by 
constructing a document-term matrix (DTM) and then multiplying the matrices 
together.

```{r}
tokens <- word_tokenizer(tidy_ref$text)
it <- itoken(tokens)
# use vocab from last GloVe model
vectorizer <- vocab_vectorizer(vocab)
dtm <- create_dtm(it, vectorizer)
tfidf <- TfIdf$new()
dtm_tfidf <- fit_transform(dtm, tfidf)
sentence_vectors <- dtm_tfidf %*% word_vectors2
```

I have to admit I'm a little surprised that worked; I was expecting that step to
be a lot harder! The next step is to revisit the multivariate normality 
assumption with this new data. There is a heap more of it (as there are more
sentences in the corpus than words in the vocabulary) so the formal tests will
definitely not be satisfied, but **ggplot2** should still be able to handle the
full dataset.

```{r}
sen_plot <- as.data.frame(as.matrix(sentence_vectors[,c(1,2,25,50)]))
names(sen_plot) <- c("dim_1", "dim_2", "dim_25", "dim_50")
ggplot(sen_plot) +
  geom_point(aes(x=dim_1, y=dim_2), alpha = 0.05) +
  geom_density2d(aes(x=dim_1, y=dim_2))
ggplot(sen_plot) +
  geom_point(aes(x=dim_25, y=dim_50), alpha = 0.05) +
  geom_density2d(aes(x=dim_25, y=dim_50))
```

These plots look close enough to multvariate normal for the purpose of this
project, so I think we can call that a win for now. The next thing to do is 
check whether the position of these sentences in the GloVe vector space has any
meaning. First we need to set up the similarity function, which will give (in 
order):

1. the sentence being compared
1. the nearest sentence
1. the second-nearest sentence
1. the furthest sentence
1. the second-furthest sentence

```{r}
dimnames(sentence_vectors) <- 
  list(str_replace_all(tidy_ref$text,"\\s+", " ") , NULL)
get_nearest_and_farthest <- function(x, vecs) {
  vecs <- as.matrix(vecs)
  sim_vec <- sim2(x = vecs, y = vecs[x,,drop=F], 
                  method = "cosine", norm = "l2")[,1] %>% 
               sort(decreasing = TRUE)
  out <- c(head(sim_vec,3),tail(sim_vec,2))
  out
}
```

We can now see if similar sentences are being discovered by the model - here 
goes nothing!

```{r}
get_nearest_and_farthest(1, sentence_vectors)
get_nearest_and_farthest(1234, sentence_vectors)
get_nearest_and_farthest(5678, sentence_vectors)
get_nearest_and_farthest(9999, sentence_vectors)
```

I don't want to get too carried away here, but it looks like it's working! The
first three sentences in each comparison are similar, and the last two are very
dissimilar! I'm going to claim that as a win for now, and I'm comfortable that 
the first assumption under test in this script is viable.
