---
title: "UK REF Impact Case Studies - Topic Modelling"
output: github_document
---
  
_This script explores the UK REF Impact Case Studies dataset. The dataset was 
previously extracted from <http://impact.ref.ac.uk/CaseStudies/> using the 
refimpact package (which was developed as part of this iLab project) and saved
locally to allow for faster access._

## Preparation

```{r, message=FALSE}
library(text2vec)
library(stringr)
```

## Data Preparation

Firstly we will load the data into memory and clean it up a little, then create
a Document Term Matrix

```{r}
ref <- readRDS("~/ref_data.rds")
prepare_text <- function(x) {
  x %>% 
  str_to_lower %>% 
  str_replace_all("[^[:alnum:]]", " ") %>% 
  str_replace_all("\\s+", " ")
}
ref$ImpactDetails <- prepare_text(ref$ImpactDetails)
it <- itoken(ref$ImpactDetails)
v <- create_vocabulary(it) %>% 
  prune_vocabulary(doc_proportion_max = 0.1, term_count_min = 5)
vectorizer <- vocab_vectorizer(v)
dtm <- create_dtm(it, vectorizer)
```

Now we will apply the TF-IDF transformation and fit a Latent Semantic Analysis
topic model. Given that there are 36 units of assessment, let's try fitting a 
model with 36 topics.

```{r}
tfidf <- TfIdf$new()
lsa <- LSA$new(n_topics = 36)
dtm_tfidf_lsa <- dtm %>% 
  fit_transform(tfidf) %>% 
  fit_transform(lsa)
head(dtm_tfidf_lsa)
```

That was quick! We could stop and visualise here, but text2vec includes a quick
visualisation method for Latent Dirichtlet Analysis (which should be a more
useful model anyway) so we might as well just skip to that.

```{r}
tokens <- ref$ImpactDetails %>% 
  prepare_text() %>% 
  word_tokenizer
it <- itoken(tokens, ids = ref$CaseStudyId)
v <- create_vocabulary(it) %>% 
  prune_vocabulary(term_count_min = 10, doc_proportion_max = 0.2)
vectorizer <- vocab_vectorizer(v)
dtm <- create_dtm(it, vectorizer, type = "lda_c")

lda_model <- LDA$new(n_topics = 36, 
                     vocabulary = v, 
                     doc_topic_prior = 0.1, 
                     topic_word_prior = 0.01)
doc_topic_distr <- lda_model$fit_transform(dtm, 
                                           n_iter = 1000, 
                                           convergence_tol = 0.01, 
                                           check_convergence_every_n = 10)

lda_model$plot()
```




